Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-03-31 document/ms.tex
adding discussion of synthetic catalogs

\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{url}
\usepackage{amssymb,amsmath}

@@

@@

@@

\begin{figure}
\begin{center}
\includegraphics{figures/separable-radius.pdf}
\end{center}
\caption{%
The inferred distribution of exoplanet radii assuming a separable rate
function.
The black lines are draws from the posterior probability distribution of bin
heights and the red points are previous measurements based on the same
dataset (\citealt{petigura}).
\figlabel{separable-radius}}
\end{figure}

@@

@@

\bibitem[Petigura \etal(2013)]{petigura}
Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013,
Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})

>>>>

\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{url}
\usepackage{amssymb,amsmath}
\usepackage{subfigure}

@@

\dfmtodo{Discuss binned model and choice of priors.}

\dfmtodo{Show where the normalization comes in and describe how it acts
qualitatively.}

@@

\section{Synthetic catalogs}

The likelihood function in \eq{true-like} provides a \emph{probability
distribution over catalogs} conditioned on the parameters of the rate
function.
Therefore, it is also the description of a \emph{generative process} for
synthetic catalogs.
To demonstrate the effectiveness of our model, we generate synthetic catalogs
from known distributions in period and radius, and test the performance of our
method on the resulting data.
For comparison, we also implement the one-over-efficiency weighting procedure
(see, for example, \citealt{howard, dressing, petigura}) and apply it to the
same datasets.

\paragraph{The generative process}
The variable rate Poisson likelihood function in \eq{true-like} can be sampled
generally for arbitrary forms of the rate function using a procedure called
``thinning'' (\citealt{poisson}).
In this \paper, however, we will only consider binned representations of
\rate, greatly simplifying the generative process.
Conditioned on a set of rate function parameters and a given completeness
function (here we'll use the injections from \citealt{petigura}), we compute
the observable rate grid from \eq{obs-rate}.
Then in each bin, we sample $J$ candidates uniformly in period and
radius (as defined by the bin edges) where $J$ is a Poisson random variable
sampled with the rate given by the observable rate in this bin.
\dfmtodo{Make this description not be the most confusing thing ever.}
Then, we add 30\% measurement uncertainties to the radius measurements where
30\% was chosen to match to properties of the real catalog
(\citealt{petigura}).

\paragraph{Results}
See \fig{synthetic-radius}.
As expected, the forward model is a much lower variance estimator.

\begin{figure}[htbp]
\begin{center}
\subfigure[one-over-efficiency]{%
\includegraphics[width=0.45\textwidth]{figures/synthetic-vmax-radius.pdf}
}
\subfigure[fully-marginalized forward model]{%
\includegraphics[width=0.45\textwidth]{figures/synthetic-radius.pdf}
}
\end{center}
\caption{%
The inferred distribution of exoplanet radii based on a noisy observation of a
synthetic catalog.
The red dotted line shows the true distribution function and the black
histograms are the constraints from the different inference techniques.
\figlabel{synthetic-radius}}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics{figures/separable-radius.pdf}
\end{center}
\caption{%
The inferred distribution of exoplanet radii assuming a separable rate
function.
The black lines are draws from the posterior probability distribution of bin
heights and the red points are previous measurements based on the same
dataset (\citealt{petigura}).
\figlabel{separable-radius}}
\end{figure}

@@

\bibitem[Dressing  \& Charbonneau(2013)]{dressing}
Dressing, C.~D., \& Charbonneau, D.\ 2013, \apj, 767, 95 (\arxiv{1302.1647})

@@

\bibitem[Howard \etal(2012)]{howard}
Howard, A.~W., Marcy, G.~W., Bryson, S.~T., et al.\ 2012, \apjs, 201, 15
(\arxiv{1103.2541})

\bibitem[Lewis \& Shedler(1979)]{poisson}
Lewis, P.~A.~W., \& Shedler, G.~S.\ 1979, Naval Research Logistics Quarterly,
26, 403

\bibitem[Petigura \etal(2013a)]{petigura-a}
Petigura, E.~A., Marcy, G.~W., \& Howard, A.~W.\ 2013a, \apj, 770, 69
(\arxiv{1304.0460})

\bibitem[Petigura \etal(2013b)]{petigura}
Petigura, E.~A., Howard, A.~W., \& Marcy, G.~W.\ 2013b,
Proceedings of the National Academy of Science, 110, 19273 (\arxiv{1311.6806})
